DATA 624 Homework 3

### 2020-02-21 DATA624/HW3.Rmd
More work on HW 3

**The outlier causes the series to be slighly larger, and the troughs to be overstated.**

@@

```{r}
plastics_with_outlier_near_the_end <- plastics
plastics_with_outlier_near_the_end[50] <- plastics_with_outlier_near_the_end[50] + 500
decompose_plastics_with_outlier_near_the_end <- decompose(plastics_with_outlier_near_the_end, type="multiplicative")
seasonally_adjusted_plastics_with_outlier_near_the_end <- plastics_with_outlier_near_the_end / decompose_plastics_with_outlier_near_the_end$seasonal

plastics_with_outlier_in_the_middle <- plastics
plastics_with_outlier_in_the_middle[30] <- plastics_with_outlier_in_the_middle[30] + 500
decompose_plastics_with_outlier_in_the_middle <- decompose(plastics_with_outlier_in_the_middle, type="multiplicative")
seasonally_adjusted_plastics_with_outlier_in_the_middle <- plastics_with_outlier_in_the_middle / decompose_plastics_with_outlier_in_the_middle$seasonal

autoplot(plastics, series = 'original data') +
  autolayer(seasonally_adjusted_plastics_with_outlier_near_the_end, series = 'with outlier near the end') +
  autolayer(seasonally_adjusted_plastics_with_outlier_in_the_middle, series = 'with outlier in the middle') +
  ylab("Sales (thousands)") +
  ggtitle("Seasonally Adjusted Plastic Product A Sales") +
  scale_color_brewer(palette = "Set1")
```

@@

**There are some spikes in the remainder early on (circa 1983) and around 2000.  I also find it interesting to see the seasonal component is decreasing over time in the X11 decomposition.**

>>>>

**The outlier causes the series to be slighly higher than the seasonally adjusted series without an outlier.  There are also some troughs in the series with an outlier that don't exist in the series without an outlier.  So the addition of the outlier causes the model to find troughs in the data it wouldn't otherwise, and slighly overstate the level of the seasonal component.**

@@

```{r}
# Near the End
plastics_with_outlier_near_the_end <- plastics
plastics_with_outlier_near_the_end[50] <- plastics_with_outlier_near_the_end[50] + 500
decompose_plastics_with_outlier_near_the_end <- decompose(plastics_with_outlier_near_the_end, type="multiplicative")
seasonally_adjusted_plastics_with_outlier_near_the_end <- plastics_with_outlier_near_the_end / decompose_plastics_with_outlier_near_the_end$seasonal

# In the Middle
plastics_with_outlier_in_the_middle <- plastics
plastics_with_outlier_in_the_middle[30] <- plastics_with_outlier_in_the_middle[30] + 500
decompose_plastics_with_outlier_in_the_middle <- decompose(plastics_with_outlier_in_the_middle, type="multiplicative")
seasonally_adjusted_plastics_with_outlier_in_the_middle <- plastics_with_outlier_in_the_middle / decompose_plastics_with_outlier_in_the_middle$seasonal

# Plot both series
autoplot(plastics, series = 'original data') +
  autolayer(seasonally_adjusted_plastics_with_outlier_near_the_end, series = 'with outlier near the end') +
  autolayer(seasonally_adjusted_plastics_with_outlier_in_the_middle, series = 'with outlier in the middle') +
  ylab("Sales (thousands)") +
  ggtitle("Seasonally Adjusted Plastic Product A Sales") +
  scale_color_brewer(palette = "Set1")
```

@@

**There are some spikes in the remainder early on (circa 1983) and around 2000.  That indicates the presense of some outliers.**
